### 2020-02-23 2020.csic5011/slides/homework01.tex
homework01

\begin{enumerate}
\item Set up data matrix $X=(x_1,\ldots,x_n)\in \R^{p\times n}$;
\item Compute the sample mean $\hat{\mu}_n$ and form $\tilde{X}= X - e \hat{\mu}_n^T$;
\item Compute top $k$ SVD of $\tilde{X} = U S_k V^T$; 
\item Plot eigenvalue curve, i.e. $i$ vs. $\lambda_i (\hat{\Sigma}_n)/tr(\hat{\Sigma}_n)$ ($i=1,\ldots,k$), with top-$k$ eigenvalue $\lambda_i$ for sample covariance matrix $\hat{\Sigma}_n=\frac{1}{n}\tilde{X}*\tilde{X}^T$, which gives you explained variation of data by principal components;
\item Use {\texttt{imshow}} to visualize the mean and top-$k$ principle components as \emph{left} singular vectors $U=[u_1,\ldots,u_k]$; 
\item For $k=1$, sort the image data $(x_i)$ ($i=1,\ldots,n$) according to the top \emph{right} singular vectors, $v_1$, in an ascending order; 
\item For $k=2$, scatter plot $(v_1,v_2)$ and select a grid on such a plane to show those images on the grid (e.g. Figure 14.23 in book [ESL]: Elements of Statistical Learning).   
\item[(h)*] You may try the parallel analysis with permutation test to see how many significant principle components you will obtain. 
\end{enumerate}

@@

>>>>

\begin{enumerate}
\item Set up data matrix $X=(x_1,\ldots,x_n)\in \R^{p\times n}$;
\item Compute the sample mean $\hat{\mu}_n$ and form $\tilde{X}= X - e \hat{\mu}_n^T$;
\item Compute top $k$ SVD of $\tilde{X} = U S_k V^T$; 
\item Plot eigenvalue curve, i.e. $i$ vs. $\lambda_i (\hat{\Sigma}_n)/tr(\hat{\Sigma}_n)$ ($i=1,\ldots,k$), with top-$k$ eigenvalue $\lambda_i$ for sample covariance matrix $\hat{\Sigma}_n=\frac{1}{n}\tilde{X}*\tilde{X}^T$, which gives you explained variation of data by principal components;
\item Use {\texttt{imshow}} to visualize the mean and top-$k$ principle components as \emph{left} singular vectors $U=[u_1,\ldots,u_k]$; 
\item For $k=1$, order the images $(x_i)$ ($i=1,\ldots,n$) according to the top first \emph{right} singular vector, $v_1$, in an ascending order of $v_1(i)$; 
\item For $k=2$, scatter plot $(v_1,v_2)$ and select a grid on such a plane to show those images on the grid (e.g. Figure 14.23 in book [ESL]: Elements of Statistical Learning).   
\item[(h)*] You may try the parallel analysis with permutation test to see how many significant principle components you will obtain. 
\end{enumerate}

@@

\item {\em Distance:} Suppose that $d: \R^p \times \R^p \to \R$ is a distance function. 
\begin{enumerate}
\item Is $d^2$ a distance function? Prove or give a counter example.
\item Is $\sqrt{d}$ a distance function? Prove or give a counter example.
\end{enumerate}
